---
title: "p8105_hw6_HL3859"
author: "Hanrui Li"
date: "2024-12-02"
output: github_document
---

# Problem 1

```{r message=FALSE}
library(p8105.datasets)
library(dplyr)
library(broom)
library(purrr)
library(ggplot2)

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
set.seed(1)

bootstrap_results =
  replicate(5000, {
    boot_sample = weather_df[sample(1:nrow(weather_df), replace = TRUE), ]
    model = lm(tmax ~ tmin, data = boot_sample)
    model_tidy = tidy(model)
    beta_0 = model_tidy$estimate[1]
    beta_1 = model_tidy$estimate[2]
    model_glance = glance(model)
    r_squared = model_glance$r.squared
    log_beta <- log(beta_0 * beta_1)
    c(r_squared = r_squared, log_beta = log_beta)
})

bootstrap_df = data.frame(t(bootstrap_results))
colnames(bootstrap_df) = c("r_squared", "log_beta")

quantile(bootstrap_df$r_squared, probs = c(0.025, 0.975))

ggplot(bootstrap_df, aes(x = r_squared)) +
  geom_density() +
  labs(title = "Distribution of Bootstrap R-squared Values", 
       x = "R-squared Estimtate", 
       y = "Density") +
  theme_minimal()
```

The distribution of R-squared values from the bootstrap samples is approximately normal, centered around 0.91. The density curve suggests that most of the estimates are concentrated near 0.91, with a slight tail extending toward lower values. This indicates that the linear model explains a majority of the variance in the relationship between tmin and tmax.

```{r}
quantile(bootstrap_df$log_beta, probs = c(0.025, 0.975))

ggplot(bootstrap_df, aes(x = log_beta)) +
  geom_density() +
  labs(title = "Distribution of Bootstrap Log(Intercept * tmin)", 
       x = "Log(Intercept * tmin)",
       y = "Density") +
  theme_minimal()
```

The distribution of the log of the product of the intercept and the tmin variable is also approximately normal, with values centered around 2.0. The peak density indicates a high concentration of estimates around this value, and the distribution has a slight tail on the higher end. This suggests the log-transformed product of the intercept and tmin is consistent across the bootstrap samples.


# Problem 2







