---
title: "p8105_hw6_HL3859"
author: "Hanrui Li"
date: "2024-12-02"
output: github_document
---

# Problem 1

```{r message=FALSE}
library(p8105.datasets)
library(tidyverse)
library(dplyr)
library(broom)
library(purrr)
library(ggplot2)

weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
set.seed(1)

bootstrap_results =
  replicate(5000, {
    boot_sample = weather_df[sample(1:nrow(weather_df), replace = TRUE), ]
    model = lm(tmax ~ tmin, data = boot_sample)
    model_tidy = tidy(model)
    beta_0 = model_tidy$estimate[1]
    beta_1 = model_tidy$estimate[2]
    model_glance = glance(model)
    r_squared = model_glance$r.squared
    log_beta <- log(beta_0 * beta_1)
    c(r_squared = r_squared, log_beta = log_beta)
})

bootstrap_df = data.frame(t(bootstrap_results))
colnames(bootstrap_df) = c("r_squared", "log_beta")

quantile(bootstrap_df$r_squared, probs = c(0.025, 0.975))

ggplot(bootstrap_df, aes(x = r_squared)) +
  geom_density() +
  labs(title = "Distribution of Bootstrap R-squared Values", 
       x = "R-squared Estimtate", 
       y = "Density") +
  theme_minimal()
```

The distribution of R-squared values from the bootstrap samples is approximately normal, centered around 0.91. The density curve suggests that most of the estimates are concentrated near 0.91, with a slight tail extending toward lower values. This indicates that the linear model explains a majority of the variance in the relationship between tmin and tmax.

```{r}
quantile(bootstrap_df$log_beta, probs = c(0.025, 0.975))

ggplot(bootstrap_df, aes(x = log_beta)) +
  geom_density() +
  labs(title = "Distribution of Bootstrap Log(Intercept * tmin)", 
       x = "Log(Intercept * tmin)",
       y = "Density") +
  theme_minimal()
```

The distribution of the log of the product of the intercept and the tmin variable is also approximately normal, with values centered around 2.0. The peak density indicates a high concentration of estimates around this value, and the distribution has a slight tail on the higher end. This suggests the log-transformed product of the intercept and tmin is consistent across the bootstrap samples.


# Problem 2

```{r}
homicide_df = 
  read.csv("homicide-data.csv", na = c("","NA","Unknown")) |>
  mutate(
    city_state = paste(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    solved = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest" ~ 0,
      disposition == "Closed by arrest" ~ 1)) |>
  select(city_state, victim_age, victim_race, victim_sex, solved) |>
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) |>
  filter(victim_race %in% c("White", "Black"))
```

```{r}
baltimore_df = 
  homicide_df |>
  filter(city_state == "Baltimore, MD")

glm_model = glm(solved ~ victim_age + victim_sex + victim_race, 
                 data = baltimore_df, 
                 family = binomial)

broom::tidy(glm_model, conf.int = TRUE) |>
  mutate(odds_ratio = exp(estimate),
         CI_lower = exp(conf.low),
         CI_upper = exp(conf.high)) |>
  select(term, estimate, odds_ratio, CI_lower, CI_upper) |>
  knitr::kable(digits = 3)
```

```{r}
get_or_ci =
  function(city_data) {
    model = glm(solved ~ victim_sex + victim_age + victim_race, 
               data = city_data, 
               family = binomial)
    broom::tidy(model, conf.int = TRUE) |>
  mutate(odds_ratio = exp(estimate),
         CI_lower = exp(conf.low),
         CI_upper = exp(conf.high)) |>
  select(term, estimate, odds_ratio, CI_lower, CI_upper)
}

city_or_ci =
  homicide_df |>
  nest(data = -city_state) |>
  mutate(model_results = map(data, get_or_ci)) |>
  filter(!map_lgl(model_results, is.null)) |>
  unnest(model_results)

city_or_ci |>
  filter(term == "victim_sexMale") |>
  arrange(odds_ratio) |>
  ggplot(aes(x = reorder(city_state, odds_ratio), y = odds_ratio)) +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  labs(title = "Adjusted Odds Ratios for Solving Homicides by City",
       x = "City, State",
       y = "Adjusted Odds Ratio") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


